---
title: "Git, GitHub and RMarkdown"
output: html_document
date: "2024-10-22"
---

#### Packages

```{r, message = FALSE}
library(tidyverse)
library(janitor)
library(sf)
library(countrycode)
```

#### Reading in

Gender inequality index csv

```{r, message=FALSE}
GII <- read_csv("data/hdr-data.csv") %>%
  clean_names()

head(GII, 10)
```

Reading in World geojson data.

Creating a iso3c column using countrycode package to be consistent with the GII data

```{r}
world_sf <- read_sf("data/World_Countries_(Generalized)_9029012925078512962.geojson") %>%
  mutate(iso3c = countrycode(ISO, origin = 'iso2c',destination = 'iso3c', nomatch = NA)) %>%
  select(iso3c)

head(world_sf)
```

#### Tidying

-   Filtering to only include relevant values (GII) and the years of interest (2010 and 2019).

-   Renaming country_iso_code to iso3c for ease of merging later with world_sf.

-   Selecting only the relevant columns.

-   Pivoting wider to convert the data to a tidy format for ease of merging.

-   Creating the difference between 2019 and 2010 column

```{r}
unique(GII$indicator) #shows unique values for indicator columns

GII_clean <- GII %>%
  filter(indicator == "Gender Inequality Index (value)",
         year == '2010' | year == '2019') %>%
  rename(iso3c = country_iso_code) %>%
  select(iso3c,country,year,value)%>%
  pivot_wider(names_from = year, values_from = value, names_prefix = 'y') %>%
  mutate(difference = y2019 - y2010)
```

```{r}
head(GII_clean)
```

#### Merging

Merging world sf and GII data using an inner join to include countries (iso3c codes) that appear in both datasets.

```{r}
GII_world <- GII_clean %>%
  inner_join(.,
        world_sf,
        by = 'iso3c')
```

```{r}
head(GII_world)
```
